Quantum Algorithms for Portflolio Optimisation and Option pricing

### 2024-06-26 project/paper-v2.tex
corrections-figures

\section{Quantum Approaches}\label{sec:literature2}
Having reviewed the classical approaches for both portfolio optimisation and option pricing, it is time to 
introduce the quantum approaches to these problems. In this section we will review the implementations of the quantum algorithms discussed in Sec. \ref{sec:background} for these tasks.
\subsection{Portfolio Optimisation}
In 2023, Buonaiuto et al. \cite{Buonaiuto2023} published a paper where they implemented the VQE algorithm in multiple simulations and in real quantum computers. They used four assets (Apple, IBM, Netflix and Tesla) 
historical data, and the MVPO model to state the optimisation problem. 
\begin{figure}[h!]
\centering 
\includegraphics[width=0.45\textwidth]{data-paper.png}
  \caption{\label{fig:data-set} Sample of the historical data of of the assets Apple, IBM, Netflix and Tesla used for the VQE implementation by 
Biunaiuto et al. \cite{Buonaiuto2023}.}
\end{figure}

They mapped the MVPO problem into a QUBO problem (Eq. \ref{eq:21} and Eq. \ref{eq:22}), and then converted it into quantum operators using 
the Ising Hamiltonian, as proposed by \cite{Tilly2022}.  Prices, returns, and covariance matrix of the assets were calculated using the historical data (see Fig. \ref{fig:data-set}). 
\begin{equation}
  \label{eq:21}
  \max_{b} \mathcal{L}(b) =\max_{b}\left(\mu''^{T} b - qb^{T} \Sigma'' b - \lambda(P''^{T}b- 1)^2\right)
\end{equation}
\begin{equation}
  \label{eq:22}
  \mathcal{H} = \sum_{i} h_i Z_i + \sum_{i,j} J_{ij} Z_i\otimes Z_j + \lambda \left(\sum_{i}\pi_i Z_i -\beta\right)^2
\end{equation}

@@

The results of the VQE implementation are shown in Fig. \ref{fig:results}. The results were contrasted with a classical solution. The Toronto, Kolkata and Auckland quantum computers performances perfectly match the classical result. Different ansatz were used for the different quantum computers. 
\begin{figure*}[t!]
\centering
\includegraphics[width=.90\textwidth]{results-vqe.png}
  \caption{\label{fig:results} Results of the VQE implementation by Buonaiuto et al. \cite{Buonaiuto2023}. The optimal asset allocation for the four assets considered is shown. Extracted from 
  \cite{Buonaiuto2023}}
\end{figure*}

@@

>>>>

\section{Quantum Approaches}\label{sec:literature2}
Having reviewed the classical approaches for both portfolio optimisation and option pricing, it is time to 
introduce the quantum approaches to these problems. In this section we will review the implementations of the quantum algorithms discussed in Sec. \ref{sec:background} for these tasks.
\subsection{Portfolio Optimisation}
In 2023, Buonaiuto et al. \cite{Buonaiuto2023} published a paper where they implemented the VQE algorithm in multiple simulations and in real quantum computers. They used four assets (Apple, IBM, Netflix and Tesla) 
historical data, and the MVPO model to state the optimisation problem.

\begin{table}[h]
    \centering
     \caption{Sample of the historical data of the assets Apple, IBM, Netflix and Tesla used for the VQE implementation by Biunaiuto et al. \cite{Buonaiuto2023}.}
    \begin{tabular}{cccccc}
        \hline
        \textbf{N.} & \textbf{Date} & \textbf{AAPL} & \textbf{IBM} & \textbf{NFLX} & \textbf{TSLA} \\
        \midrule
        0  & 2011-12-23 & 12.260060 & 116.121811 & 10.374286 & 1.860000 \\
        1  & 2011-12-27 & 12.357327 & 116.247513 & 10.085714 & 1.904667 \\
        2  & 2011-12-28 & 12.239082 & 115.644180 & 9.885714 & 1.900667 \\
        3  & 2011-12-29 & 12.314467 & 117.020638 & 9.900000 & 1.915333 \\
        4  & 2011-12-30 & 12.310820 & 115.575020 & 9.898571 & 1.904000 \\
        5  & 2012-01-03 & 12.500192 & 117.096062 & 10.320000 & 1.872000 \\
        6  & 2012-01-04 & 12.567370 & 116.618378 & 11.492857 & 1.847333 \\
        7  & 2012-01-05 & 12.706894 & 116.065247 & 11.328571 & 1.808000 \\
        8  & 2012-01-06 & 12.839729 & 114.732758 & 12.327143 & 1.794000 \\
        9  & 2012-01-09 & 12.819363 & 114.135643 & 14.025714 & 1.816667 \\
        \hline
    \end{tabular} 
    \label{tab:stock_prices}
\end{table}

They mapped the MVPO problem into a QUBO problem (Eq. \ref{eq:21} and Eq. \ref{eq:22}), and then converted it into quantum operators using 
the Ising Hamiltonian, as proposed by \cite{Tilly2022}.  Prices, returns, and covariance matrix of the assets were calculated using the historical data (see Table \ref{tab:stock_prices}). 
\begin{equation}
  \label{eq:21}
  \max_{b} \mathcal{L}(b) =\max_{b}\left(\mu''^{T} b - qb^{T} \Sigma'' b - \lambda(P''^{T}b- 1)^2\right)
\end{equation}
\begin{equation}
  \label{eq:22}
  \mathcal{H} = \sum_{i} h_i Z_i + \sum_{i,j} J_{ij} Z_i\otimes Z_j + \lambda \left(\sum_{i}\pi_i Z_i -\beta\right)^2
\end{equation}

@@

The results of the VQE implementation are shown in Fig. \ref{fig:results}. The results were contrasted with a classical solution. The Toronto, Kolkata and Auckland quantum computers performances perfectly match the classical result. Different ansatz were used for the different quantum computers. 
\begin{figure*}[t!]
\centering
\includegraphics[width=.90\textwidth]{vqe-results.pdf}
  \caption{\label{fig:results} Results of the VQE implementation by Buonaiuto et al. \cite{Buonaiuto2023}. The blue line represents the trade of curve between the expected return and the volatility. Adapted from 
  \cite{Buonaiuto2023}}
\end{figure*}

The results show that the performance of the algorithm is highly dependent on the ansatz selected and the hardware topology, which dictates the physical arrangement of the qubits and thus, the level of efficiency  the state preparation can be performed with.

Furthermore, the authors mention that a bigger quantum volume is needed to perform deeper circuits without  an exponential 
increase in the error rate.

@@

Future research in NISQ devices with larger quantum volumes, lower error rates, and error mitigation protocols would be crucial 
to further develop these quantum algorithms.
